---
title: "Fit bivariate Gaussian copula to La and Ni data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####Fitting bivariate copula to Olympic Dam data. Choice of pairs of elements was made on the basis of scatterplotof La and Ni.
```{r include = FALSE}
library(compositions)
library(copula)
library(VineCopula)
library(dplyr)
```

Read data in, filter out samples from drillhole 1988 and perform log transformation

```{r}
hem <- read.csv("hem_sorted.txt", header=TRUE, sep="\t")
hem <- hem %>% filter(Drillhole != "RD1988")
hem.log <- log1p(hem[,8:53])
```

Produce a scatterplot of La vs Ni and calculate Pearson's correlation coefficient
```{r}
plot(hem.log$La139,hem.log$Ni60)
abline(lm(hem.log$La139~hem.log$Ni60),col='red',lwd=1)
cor(hem.log$La139,hem.log$Ni60,method='spearman')
```


Create normal two-dimensional copula
```{r}
n.cop <- normalCopula(dim=2)
```

Fit copula
```{r}
set.seed(500)
m <-pobs(as.matrix(cbind(hem.log$La139,hem.log$Ni60)))
fit <- fitCopula(n.cop,m,method='ml')
rho = coef(fit)[1]
print(rho)
```

A bivariate Gaussian copula with given parameter rho
```{r}
persp(normalCopula(dim=2,rho),dCopula)
```

Sample 380 points from normal bivariate copula with a givenparameter rho
```{r}
u <- rCopula(380, normalCopula(dim=2,rho))
plot(u[,1],u[,2], col="darkblue")
cor(u, method='pearson')
```

Model the margnals: calculate mean and standard deviations for La and Ni
```{r}
La_mu <- mean(hem.log$La139)
La_sd <- sd(hem.log$La139)
Ni_mu <- mean(hem.log$Ni60)
Ni_sd <- sd(hem.log$Ni60)
```

Plot histograms of La and Ni with outlined normal distribution density function
```{r}
hist(hem.log$La139, density=20, breaks=10, prob=TRUE, 
     xlab="La", ylim=c(0, 0.7), main="normal curve over histogram")
curve(dnorm(x, mean=La_mu, sd=La_sd), col="darkblue", lwd=2, add=TRUE, yaxt="n")

hist(hem.log$Ni60, density=20, breaks=10, prob=TRUE, 
     xlab="Ni", ylim=c(0, 0.8),main="normal curve over histogram")
curve(dnorm(x, mean=Ni_mu, sd=Ni_sd), col="darkblue", lwd=2, add=TRUE, yaxt="n")

```

Simulate data with two normally distributed marginals
```{r}
copula_dist <- mvdc(copula=normalCopula(rho,dim=2), margins=c("norm","norm"),
               paramMargins=list(list(mean=La_mu, sd=La_sd),list(mean=Ni_mu, sd=Ni_mu)))

sim <- rMvdc(380,copula_dist)
```

Plot observed and simulated points
```{r}
plot(hem.log$La139,hem.log$Ni60,main='observed')
points(sim[,1],sim[,2],col='red')
legend('bottomright',c('Observed','Simulated'),col=c('black','red'),pch=21)
```